---
title: "CBRUVS Manuscript"
author: "Allyson"
date: "`r Sys.Date()`"
output: html_document
---

# Setup
```{r setup}
library(tidyverse)
library(readxl)
library(RColorBrewer)
library(patchwork)
library(scales)
library(sf)
library(ggrepel)
library(dunn.test)
library(car)
library(MASS)
library(survival)
library(plotrix)
library(ggh4x)
library(meta)
library(metafor)
library(patchwork)
library(ggspatial)

dir.create("Figures", showWarnings = FALSE)
dir.create("output_tables", showWarnings = FALSE)
```

# Site Map
```{r}
fl_shore_path <- "../../../../../Florida_Shoreline_(1_to_12%2C000_Scale)/Florida_Shoreline_(1_to_12%2C000_Scale).shp"

site_map <- read_xlsx("raw_data/site_coords.xlsx")

if (file.exists(fl_shore_path)) {
  fl_shore <- st_read(fl_shore_path, quiet = TRUE) %>% st_transform(crs = 4326)
  has_shore <- TRUE
} else {
  message("Shoreline shapefile not found.")
  has_shore <- FALSE
}

map_theme <- theme_bw() +
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "aliceblue"),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
        plot.title  = element_text(size = 11, face = "bold", hjust = 0.5))

make_map <- function(data, title, xlim, ylim) {
  p <- ggplot()
  if (has_shore) p <- p +
    geom_sf(data = fl_shore, fill = "lightgray", color = "darkgray", linewidth = 0.15)
  p +
    geom_point(data = data,
               aes(x = Longitude, y = Latitude, fill = SiteName),
               shape = 21, colour = "black", stroke = 0.4, alpha = 0.9) +
    geom_label_repel(data = data,
                     aes(x = Longitude, y = Latitude, label = SiteName),
                     size = 2.3, box.padding = 0.6, point.padding = 0.3,
                     max.overlaps = Inf, force = 12,
                     segment.size = 0.3, segment.color = "grey40",
                     label.padding = unit(0.15, "lines"), min.segment.length = 0) +
    coord_sf(xlim = xlim, ylim = ylim) +
    annotation_north_arrow(location = "tr", style = north_arrow_fancy_orienteering()) +
    annotation_scale(location = "bl") +
    labs(title = title, x = "Longitude", y = "Latitude") +
    map_theme +
    guides(size = guide_legend(override.aes = list(shape = 21, fill = "gray60"))) 
}

miami_map <- site_map %>%
  filter(Latitude >= 25.5 & Latitude <= 25.85, Longitude >= -80.25 & Longitude <= -79.95)

p5b <- make_map(miami_map, "C-BRUVS Deployments and In Situ Coral Nursery",
                xlim = c(-80.25, -79.95), ylim = c(25.5, 25.85))

ggsave("Figures/UM_Fig5_Map.pdf", p5_combined, width = 16, height = 9)
```


# Common-name lookup (used for every panel)
```{r common-names}
# Single source of truth for scientific -> common name, so all 4 panels stay
# consistent. Add new species here if they show up elsewhere.
common_names <- c(
  "Sparisoma aurofrenatum" = "Redband parrotfish",
  "Sparisoma viride"       = "Stoplight parrotfish",
  "Sparisoma rubripinne"   = "Yellowtail parrotfish",
  "Sparisoma chrysopterum" = "Redtail parrotfish",
  "Scarus iseri"           = "Striped parrotfish",
  "Scarus taeniopterus"    = "Princess parrotfish",
  "Scarus vetula"          = "Queen parrotfish",
  "Scarus guacamaia"       = "Rainbow parrotfish",
  "Scarus coelestinus"     = "Midnight parrotfish",
  "Scarus coeruleus"       = "Blue parrotfish",
  "Chaetodon capistratus"  = "Foureye butterflyfish",
  "Chaetodon striatus"     = "Banded butterflyfish",
  "Chaetodon ocellatus"    = "Spotfin butterflyfish",
  "Stegastes diencaeus"    = "Longfin damselfish",
  "Stegastes planifrons"   = "Threespot damselfish",
  "Acanthurus bahianus"    = "Ocean surgeonfish",
  "Acanthurus coeruleus"   = "Blue tang",
  "Acanthurus chirurgus"   = "Doctorfish",
  "Acanthurus sp."         = "Surgeonfish (unidentified)"
)

to_common <- function(x) {
  out <- common_names[x]
  ifelse(is.na(out), x, out)  # fall back to original if not in lookup
}
```

# Load clean data

# All data cleaning (standardizing species names, fixing typos, converting
# coral codes to scientific names, correcting known data-entry errors, etc.)
# is done once in 01_clean_data.R and documented there. This script reads
# the resulting tidy CSVs directly.
```{r load-data}
CBRUV_updated <- read_csv("data_clean/CBRUVS_bites_clean.csv", show_col_types = FALSE)
RVC_Fish_Survey <- read_csv("data_clean/RVC_fish_survey_clean.csv", show_col_types = FALSE)
PT1_clean <- read_csv("data_clean/PT1_coral_fragments_clean.csv", show_col_types = FALSE)

corallivore_species <- c("Sparisoma aurofrenatum", "Sparisoma viride", "Chaetodon capistratus")

parrotfish_corallivore_species <- c("Sparisoma aurofrenatum", "Sparisoma viride")

corallivore_species_rvc <- c(corallivore_species,
                              "Scarus vetula", "Scarus guacamaia",
                               "Scarus taeniopterus")

species_colors <- setNames(
  RColorBrewer::brewer.pal(7, "Set2"),
  c("Redband parrotfish", "Stoplight parrotfish", "Foureye butterflyfish",
    "Queen parrotfish", "Rainbow parrotfish", "Princess parrotfish", "Striped parrotfish")
)

fragment_availability_day0 <- PT1_clean %>%
  filter(Days == 0) %>%
  count(Coral, name = "available")

fragment_days_tracked <- PT1_clean %>%
  filter(Days %in% c(0, 1, 2, 7, 14)) %>%
  filter(!is.na(Mortality)) %>%
  count(Tray, Coral, name = "fragment_days")
```


### Total number of bites
```{r}
video_duration_by_day <- CBRUV_updated %>%
  distinct(Tag, Days, Video, video_duration_seconds) %>%
  group_by(Days) %>%
  summarise(total_hours = sum(video_duration_seconds, na.rm = TRUE) / 3600, .groups = "drop")

video_duration_by_day_tray <- CBRUV_updated %>%
  distinct(Tag, Days, Video, video_duration_seconds) %>%
  group_by(Days,Tag) %>%
  summarise(total_hours = sum(video_duration_seconds, na.rm = TRUE) / 3600, .groups = "drop")

panel_A_data <- CBRUV_updated %>%
  filter(Fish %in% corallivore_species) %>%
  group_by(Days, Fish) %>%
  summarise(Bites = sum(Bites), .groups = "drop") %>%
  mutate(Fish_common = to_common(Fish))

panel_A_data %>%
  group_by(Fish_common) %>%
  summarise(Bites = sum(Bites), .groups = "drop") %>%
  mutate(percent = 100 * Bites / sum(Bites))

# bar-top position + duration label for each day, used by geom_text below
total_bites_by_day <- panel_A_data %>%
  group_by(Days) %>%
  summarise(total_bites = sum(Bites), .groups = "drop") %>%
  left_join(video_duration_by_day, by = "Days")

fig2A <- panel_A_data %>%
  ggplot(aes(x = factor(Days), y = Bites, fill = Fish_common)) +
  geom_col(color = "black", width = 0.7) +
  geom_text(
    data = total_bites_by_day,
    aes(x = factor(Days), y = total_bites, label = paste0(round(total_hours, 1), " hr")),
    inherit.aes = FALSE, vjust = -0.5, size = 3
  ) +
  scale_fill_manual(values = species_colors, name = "Fish species") +
  labs(x = "Day", y = "Total bites") +
  theme_bw() +
  theme(panel.grid.minor = element_blank()) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, max(total_bites_by_day$total_bites) * 1.1))
```



### Mean C-BRUVS bite rate over time
```{r}
video_duration_lookup <- CBRUV_updated %>%
  distinct(Tag, Days, Video, video_duration_seconds)

tag_day_rates <- CBRUV_updated %>%
  group_by(Tag, Days, Fish) %>%
  summarise(total_bites = sum(Bites, na.rm = TRUE), .groups = "drop") %>%
  left_join(
    video_duration_lookup %>%
      group_by(Tag, Days) %>%
      summarise(total_minutes = sum(video_duration_seconds, na.rm = TRUE) / 60, .groups = "drop"),
    by = c("Tag", "Days")
  ) %>%
  mutate(bite_rate_per_min = total_bites / total_minutes)

panel_B_data <- tag_day_rates %>%
  filter(Fish %in% corallivore_species) %>%
  group_by(Fish, Days) %>%
  summarise(
    mean_bite_rate_per_min = mean(bite_rate_per_min, na.rm = TRUE),
    se_bite_rate_per_min = sd(bite_rate_per_min, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  mutate(Fish_common = to_common(Fish))

fig2B <- panel_B_data %>%
  ggplot(aes(x = factor(Days), y = mean_bite_rate_per_min, color = Fish_common)) +
  geom_point(position = position_dodge(width = 0.5), size = 2) +
  geom_errorbar(aes(ymin = mean_bite_rate_per_min - se_bite_rate_per_min,
                     ymax = mean_bite_rate_per_min + se_bite_rate_per_min),
                width = 0.4, linewidth = 0.8, position = position_dodge(width = 0.5)) +
  scale_color_manual(values = species_colors, name = "Fish species") +
  labs(x = "Day", y = "Mean bite rate (bites/min)") +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```


### Parrotfish age-phase composition (Initial vs Terminal) — stat only, not a figure panel
```{r agephase-stat}
agephase_stat <- CBRUV_updated %>%
  filter(Fish %in% parrotfish_corallivore_species, AgePhase %in% c("Initial", "Terminal", "Juvenile")) %>%
  count(AgePhase, Fish, name = "n") %>%
  mutate(percent = 100 * n / sum(n))

agephase_stat
```

### Parrotfish size class vs. total bites
```{r}
panel_C_data <- CBRUV_updated %>%
  filter(Fish %in% parrotfish_corallivore_species) %>%
  mutate(
    size_class = cut(Length,
                      breaks = c(0, 5, 10, 15, 20, Inf),
                      labels = c("<5 cm", "5-10 cm", "10-15 cm", "15-20 cm", ">20 cm"),
                      right = TRUE),
    Fish_common = to_common(Fish)
  ) %>%
  filter(!is.na(size_class)) %>%
  group_by(Tag, Fish_common, size_class) %>%
  summarise(total_bites = sum(Bites), .groups = "drop") %>%
  complete(Tag = unique(CBRUV_updated$Tag), Fish_common, size_class, fill = list(total_bites = 0)) %>%
  group_by(Fish_common, size_class) %>%
  summarise(
    mean_bites = mean(total_bites, na.rm = TRUE),
    se_bites = sd(total_bites, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

fig2C <- panel_C_data %>%
  ggplot(aes(x = size_class, y = mean_bites, color = Fish_common)) +
  geom_point(position = position_dodge(width = 0.5), size = 2) +
  geom_errorbar(aes(ymin = pmax(mean_bites - se_bites, 0), ymax = mean_bites + se_bites),
                width = 0.4, linewidth = 0.8, position = position_dodge(width = 0.5)) +
  scale_color_manual(values = species_colors, name = "Fish species") +
  labs(x = "Size class", y = "Mean bites per C-BRUVS (+/- SE)") +
  theme_bw() +
  theme(panel.grid.minor = element_blank())
```

### RVC fish observed by day, Days 0–14
```{r}
rvc_fish <- RVC_Fish_Survey %>%
  filter(Fish %in% corallivore_species_rvc) %>%
  mutate(Fish_common = to_common(Fish)) %>%
  group_by(Days, Fish_common) %>%
  summarise(N = sum(N), .groups = "drop")

fig2D <- rvc_fish %>%
  ggplot(aes(x = Fish_common, y = N, fill = Fish_common)) +
  geom_col(color = "black", width = 0.7) +
  facet_wrap(~ Days, nrow = 1, labeller = labeller(Days = function(x) paste("Day", x)), scales = "free_x") +
  scale_fill_manual(values = species_colors, name = "Fish species") +
  labs(x = NULL, y = "Number of individuals observed (Diver survey)") +
  theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 9),
    legend.position = "right"
  )+
  scale_y_continuous(expand = c(0, 0), limits = c(0, max(rvc_fish$N) * 1.1))
```

### Combine into Figure 2
```{r fig2-combined, fig.width=12, fig.height=12}
figure2 <- (fig2A + fig2C) / fig2B / fig2D +
  plot_annotation(tag_levels = "A")

figure2

ggsave("Figures/Figure2_combined.png", figure2, width = 12, height = 12, dpi = 300)
ggsave("Figures/Figure2_combined.pdf", figure2, width = 12, height = 12)
```

# Figure 3

### Cumulative predation prevalence
```{r}
pooled_prevalence <- PT1_clean %>%
  filter(Species != "APAL", Species != "ACER") %>%  # exclude acroporids, consistent with the GLMM
  mutate(
    Predation = if_else(Tray == "B091" & Days == 14 & UniqueCode == 127, "N", Predation),
    Cumulative = if_else(Tray == "B091" & Days == 14 & UniqueCode == 127, "N", Cumulative)
  ) %>% 
  group_by(Days, Cumulative) %>% 
  summarise(count = n()) %>% 
  mutate(Cumulative = ifelse(
    cumsum(ifelse(Cumulative == "Y" & !is.na(Cumulative) & Cumulative != "", 1, 0)) > 0, 
    "1", 
    "0")) %>% 
  pivot_wider(names_from = "Cumulative", values_from = "count") %>% 
  ungroup() %>% 
  mutate(across(c(`0`, `1`), ~ replace_na(.x, 0))) %>% 
  mutate(total = `0` + `1`) %>% 
  mutate(proportion_predation = `1`/total)

p1<-PT1_clean %>% 
  filter(Species != "APAL" & Species != "ACER") %>% 
  filter(!(Days == 0 & Predation == "Y")) %>%  # 3 corals were reported with predation
  group_by(Days, Cumulative, Coral) %>% 
  summarise(count = n()) %>% 
  mutate(Cumulative = ifelse(
    cumsum(ifelse(Cumulative == "Y" & !is.na(Cumulative) & Cumulative != "", 1, 0)) > 0, 
    "1", 
    "0")) %>% 
  pivot_wider(names_from = "Cumulative", values_from = "count") %>% 
  ungroup() %>% 
  mutate(across(c(`0`, `1`), ~ replace_na(.x, 0))) %>% 
  mutate(total = `0` + `1`) %>% 
  mutate(proportion_predation = `1`/total) %>% 
  ggplot(., aes(x=Days, y=proportion_predation, color = Coral)) +
  geom_point(alpha = 0.15) +
  geom_line(data = pooled_prevalence, aes(x = Days, y = proportion_predation),
            inherit.aes = FALSE, color = "black", linewidth = 1.1, alpha = 0.7) +
  geom_point(data = pooled_prevalence, aes(x = Days, y = proportion_predation),
             inherit.aes = FALSE, color = "black", size = 2, alpha = 0.5) +
  geom_line() +
  theme_bw() +
  ylab("Cumulative Predation Prevalence") +
  xlab("Days Since Deployment") +
  theme(
    legend.text = element_text(face = "italic"),
    legend.position = "none"
  ) + scale_x_continuous(breaks = c(0, 1, 2, 7, 14, 21, 28, 35, 42),
                      limits = c(0, 42),
                      expand = c(0, 0))
  

p1
#ggsave("Cumulative_Predation_Prevalence_allcorals.pdf", width = 8, height = 6)
```


### Predation intensity (tissue mortality) per coral species, accounting for dropout
```{r}
fragment_trajectories <- PT1_clean %>%
filter(Species != "APAL" & Species != "ACER") %>% 
  mutate(Mortality = as.numeric(Mortality)) %>%
  filter(!is.na(Mortality)) %>%
  group_by(Tray, UniqueCode, Coral) %>%
  summarise(
    ever_predated = any(Predation == "Y", na.rm = TRUE),
    max_mortality = if_else(ever_predated, max(Mortality), 0),
    reached_100 = if_else(ever_predated, any(Mortality == 100), FALSE),
    reached_30_cutoff = if_else(ever_predated, max_mortality >= 30, FALSE),
    day_of_endpoint = if_else(reached_100 | reached_30_cutoff, max(Days), NA_real_),
    .groups = "drop"
  )

n_deployed <- PT1_clean %>%
  filter(!is.na(Mortality)) %>%
  distinct(Tray, UniqueCode, Coral) %>%
  count(Coral, name = "n_fragments_deployed")

panel_X_data <- fragment_trajectories %>%
  group_by(Coral) %>%
  summarise(
    n_predated = sum(ever_predated),
    mean_max_mortality = mean(max_mortality[ever_predated]),
    se_max_mortality = sd(max_mortality[ever_predated]) / sqrt(sum(ever_predated)),
    pct_died_100 = 100 * mean(reached_100[ever_predated]),
    pct_censored_30 = 100 * mean(reached_30_cutoff[ever_predated]),
    mean_day_to_endpoint = mean(day_of_endpoint, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  complete(Coral = n_deployed$Coral,
           fill = list(n_predated = 0, mean_max_mortality = 0, se_max_mortality = 0,
                       pct_died_100 = 0, pct_censored_30 = 0, mean_day_to_endpoint = NA_real_)) %>%
  left_join(n_deployed, by = "Coral") %>%
  mutate(pct_predated = round(100 * n_predated / n_fragments_deployed, 1)) %>%
  relocate(n_fragments_deployed, n_predated, pct_predated, .after = Coral) %>%
  arrange(desc(mean_max_mortality))

write_csv(panel_X_data, "output_tables/Figure3_predation_data.csv")

p2 <- panel_X_data %>%
  filter(Coral != "Acropora palmata" & Coral != "Acropora cervicornis") %>%
  ggplot(aes(x = fct_reorder(Coral, mean_max_mortality), y = mean_max_mortality, fill = Coral)) +
  geom_col(color = "black", width = 0.7) +
  geom_errorbar(aes(ymin = mean_max_mortality - se_max_mortality,
                     ymax = mean_max_mortality + se_max_mortality),
                width = 0.3) +
  labs(x = "Coral species", y = "Mean maximum tissue mortality (%, +/- SE)") +
  coord_flip() +
  theme_bw() +
  theme(panel.grid.minor = element_blank(), axis.text.y = element_text(face = "italic"), legend.position = "none") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 60)) 

p2
```

```{r}
cowplot::plot_grid(p1, p2, ncol = 2, labels = c("A", "B"), label_size = 12)
ggsave("Figures/Figure3_predation.png", width = 12, height = 6, dpi = 300)
```


# Figure 4

### Fish x coral selectivity index, day- and tray-matched, with significance test
```{r}
# alpha separately per Tray x Day (replicate = Tray-Day,same logic as Burkepile et al. 2019's per-individual-follow electivity values), then running a one-sample t-test against 1/n_corals for each cell, matching their statistical approach (Manly's alpha + one-sample t-test).

availability_by_tray_day <- PT1_clean %>%
  filter(!is.na(Mortality)) %>%  # still actively tracked as of this day
  count(Tray, Days, Coral, name = "available")

all_coral_species <- unique(PT1_clean$Coral)
all_trays <- unique(PT1_clean$Tray)
all_days_bruvs <- c(0, 1, 2, 7, 14)

bites_by_tray_day <- CBRUV_updated %>%
  filter(Fish %in% corallivore_species) %>%
  group_by(Tag, Days, Fish, Coral) %>%
  summarise(Bites = sum(Bites), .groups = "drop") %>%
  rename(Tray = Tag) %>%
  complete(Tray = all_trays, Days = all_days_bruvs, Fish = corallivore_species,
           Coral = all_coral_species, fill = list(Bites = 0))

alpha_by_tray_day <- bites_by_tray_day %>%
  left_join(availability_by_tray_day, by = c("Tray", "Days", "Coral")) %>%
  mutate(available = replace_na(available, 0)) %>%
  group_by(Tray, Days, Fish) %>%
  filter(sum(Bites) > 0, sum(available) > 0) %>%  # need real bites and real availability to compute a composition
  mutate(
    p_bite = Bites / sum(Bites),
    p_avail = available / sum(available),
    raw_selectivity = if_else(p_avail > 0, p_bite / p_avail, NA_real_),
    alpha = raw_selectivity / sum(raw_selectivity, na.rm = TRUE)
  ) %>%
  ungroup()

n_corals_in_panel <- length(all_coral_species)
no_preference_value <- 1 / n_corals_in_panel

# One-sample t-test per Fish x Coral cell against the no-preference value
panel_4_data <- alpha_by_tray_day %>%
  filter(!is.na(alpha)) %>%
  group_by(Fish, Coral) %>%
  summarise(
    mean_alpha = mean(alpha),
    n_replicates = n(),
    p_value = if (n() >= 2 && sd(alpha) > 0) t.test(alpha, mu = no_preference_value)$p.value else NA_real_,
    .groups = "drop"
  ) %>%
  mutate(
    significant = !is.na(p_value) & p_value < 0.05,
    Fish_common = to_common(Fish),
    Fish_common = factor(Fish_common, levels = sort(unique(Fish_common), decreasing = TRUE))
  )

panel_4_data %>% 
  arrange(Fish_common, desc(mean_alpha)) %>% 
  drop_na() %>%
  write_csv("output_tables/selectivity_data.csv")

figure4 <- panel_4_data %>%
  ggplot(aes(x = Coral, y = Fish_common, fill = mean_alpha)) +
  geom_tile(color = "black", width = 0.9) +
  geom_text(aes(label = if_else(significant, "*", "")), size = 5, vjust = 0.75) +
  scale_fill_gradient2(
    low = "firebrick", mid = "white", high = "darkblue",
    midpoint = no_preference_value,
    name = "Selectivity\n(mean alpha)"
  ) +
  labs(x = "Coral species", y = NULL) + 
       #caption = paste0("White ~ ", round(no_preference_value, 3), " (no preference); blue = preferred; red = avoided\n",
                        # "* = one-sample t-test vs. no-preference, p < 0.05 (n varies by cell -- see printed table)")) +
  theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, face = "italic", size = 9)
  )

figure4

ggsave("Figures/Figure4_selectivity.png", figure4, width = 8, height = 5, dpi = 300)
ggsave("Figures/Figure4_selectivity.pdf", figure4, width = 8, height = 5)
```

# Figure 5

### Linking individual fish bites to tissue loss at the next monitored timepoint
```{r fig5}
# linking bites from C-BRUVS to the same fragment's change in tissue mortality between day of bite and the next monitoring time point

next_day_lookup <- c(`0` = 1, `1` = 2, `2` = 7, `7` = 14, `14`=21)

genotype_fixes <- c("PLCI04" = "PCLI04", "OFAV5" = "OFAV05", "PSTRI01" = "PSTR01")

ambiguous_genotype_trays <- PT1_clean %>%
  filter(Days == 0) %>%
  count(Tray, Genotype) %>%
  filter(n > 1) %>%
  dplyr::select(Tray, Genotype)

bites_before_ambiguous_exclusion <- CBRUV_updated %>%
  filter(Fish %in% corallivore_species) %>%
  mutate(Genotype = if_else(Genotype %in% names(genotype_fixes), genotype_fixes[Genotype], Genotype)) %>%
  rename(Tray = Tag)
n_ambiguous_excluded <- bites_before_ambiguous_exclusion %>%
  semi_join(ambiguous_genotype_trays, by = c("Tray", "Genotype")) %>%
  nrow()

bite_fragment_link <- bites_before_ambiguous_exclusion %>%
  anti_join(ambiguous_genotype_trays, by = c("Tray", "Genotype")) %>%
  group_by(Tray, Days, Genotype, Fish) %>%
  summarise(total_bites = sum(Bites), .groups = "drop") %>%
  mutate(next_day = next_day_lookup[as.character(Days)])

n_unmatched <- bite_fragment_link %>%
  anti_join(PT1_clean %>% distinct(Tray, Genotype), by = c("Tray", "Genotype")) %>%
  nrow()
cat("Bite records dropped due to no genotype match in PT1:", n_unmatched, "of", nrow(bite_fragment_link), "\n")
cat("Bite records dropped due to ambiguous duplicate-genotype-within-tray fragments:", n_ambiguous_excluded,
    "of", nrow(bites_before_ambiguous_exclusion), "\n")

mortality_lookup <- PT1_clean %>%
  mutate(Mortality = as.numeric(Mortality)) %>%
  dplyr::select(Tray, Genotype, Days, Mortality)

panel_5_data <- bite_fragment_link %>%
  inner_join(mortality_lookup, by = c("Tray", "Genotype", "Days")) %>%
  rename(mortality_at_bite = Mortality) %>%
  inner_join(
    mortality_lookup %>% rename(next_day = Days, mortality_at_next = Mortality),
    by = c("Tray", "Genotype", "next_day")
  ) %>%
  filter(!is.na(mortality_at_bite), !is.na(mortality_at_next)) %>%
  mutate(Fish_common = to_common(Fish))

#does bite count predict the outcome?
panel_5_data %>%
  group_by(Fish_common) %>%
  group_modify(~ {
    if (nrow(.x) < 4) return(tibble(bites_coef = NA, bites_p = NA, n = nrow(.x)))
    m <- lm(mortality_at_next ~ total_bites + mortality_at_bite, data = .x)
    s <- summary(m)$coefficients
    tibble(bites_coef = s["total_bites", "Estimate"], bites_p = s["total_bites", "Pr(>|t|)"], n = nrow(.x))
  }) %>%
  print()

panel_5_data %>%
  group_by(Fish_common) %>%
  summarise(n_bite_events = n(), mean_bites_per_event = mean(total_bites), .groups = "drop")  # for Results

figure5 <- panel_5_data %>%
  ggplot(aes(x = total_bites, y = mortality_at_next, color = mortality_at_bite)) +
  geom_point(alpha = 0.7, size = 2) +
  geom_smooth(method = "lm", se = TRUE, color = "black", linewidth = 0.6) +
  facet_wrap(~Fish_common, scales = "free_x") +
  scale_color_viridis_c(name = "Tissue mortality\nat time of bite (%)") +
  labs(x = "Bites recorded on C-BRUVS (this timepoint)",
       y = "Tissue mortality at next monitored timepoint (%)") +
  theme_bw() +
  theme(panel.grid.minor = element_blank())

figure5

ggsave("Figures/Figure5_bite_outcome.png", figure5, width = 11, height = 4.5, dpi = 300)
ggsave("Figures/Figure5_bite_outcome.pdf", figure5, width = 11, height = 4.5)


# linear model with covariate: mortality_at_next ~ total_bites + mortality_at_bite
# Starting tissue loss included as covariate to avoid the ceiling-effect
model_results <- panel_5_data %>%
  group_by(Fish_common) %>%
  group_modify(~ {
    if (nrow(.x) < 4) return(tibble(
      n = nrow(.x), bites_coef = NA, bites_se = NA,
      bites_t = NA, bites_p = NA, r_squared = NA, model_p = NA))
    m <- lm(mortality_at_next ~ total_bites + mortality_at_bite, data = .x)
    s <- summary(m)
    coefs <- s$coefficients
    tibble(
      n = nrow(.x),
      bites_coef  = coefs["total_bites", "Estimate"],
      bites_se    = coefs["total_bites", "Std. Error"],
      bites_t     = coefs["total_bites", "t value"],
      bites_p     = coefs["total_bites", "Pr(>|t|)"],
      r_squared   = s$r.squared,
      model_p     = pf(s$fstatistic[1], s$fstatistic[2], s$fstatistic[3], lower.tail = FALSE)
    )
  })

print(model_results, width = Inf)
```


# summary tables for manuscript
```{r}
fragment_bites <- CBRUV_updated %>%
   filter(Coral != "Acropora palmata" & Coral != "Acropora cervicornis") %>%
  group_by(Tag, Genotype, Coral) %>%
  summarise(total_bites_this_fragment = sum(Bites), .groups = "drop")

grand_total_bites <- sum(fragment_bites$total_bites_this_fragment) # 460 total bites

species_bite_summary <- fragment_bites %>%
  filter(Coral != "Acropora palmata" & Coral != "Acropora cervicornis") %>%
  group_by(Coral) %>%
  summarise(
    n_fragments_bitten = n(),
    total_bites = sum(total_bites_this_fragment),
    mean_bites_per_bitten_fragment = mean(total_bites_this_fragment),
    sd_bites_per_bitten_fragment = sd(total_bites_this_fragment),
    .groups = "drop"
  ) %>%
  mutate(pct_of_total_bites = round(100 * total_bites / grand_total_bites, 2)) %>%
  left_join(fragment_availability_day0, by = "Coral") %>%
  mutate(
    pct_fragments_bitten = round(100 * n_fragments_bitten / available, 2),
    bites_display = sprintf("%.2f (%.2f)", mean_bites_per_bitten_fragment, sd_bites_per_bitten_fragment)
  ) 

write_csv(species_bite_summary, "output_tables/fragment_bites_summary.csv")
```


# summary of fish (C-BRUVS and RVC)
```{r}
RVC_Fish_Survey %>%
  filter(Fish %in% corallivore_species_rvc) %>%
  mutate(Fish_common = to_common(Fish))  %>% 
  dplyr::select(Days, N, Fish_common, `Average Size`, Min, Max) %>% 
  rename(N_RVC = N, Average_Size_RVC = `Average Size`) %>% 
  mutate(Days = as.double(Days)) %>% 
  mutate(
  Average_Size_RVC = as.numeric(Average_Size_RVC),
  Min = as.numeric(Min),
  Max = as.numeric(Max),
  size_display_RVC = if_else(
    is.na(Min) | is.na(Max) | Min == Max,
    sprintf("%.0f", Average_Size_RVC),
    sprintf("%.0f (%.0f–%.0f)", Average_Size_RVC, Min, Max)
  )
) -> RVC_Fish_Survey_subset

unique(RVC_Fish_Survey_subset$Fish_common)

CBRUV_updated %>%
  filter(Fish %in% corallivore_species_rvc) %>%
  mutate(Fish_common = to_common(Fish)) %>% 
  dplyr::select(Fish_common, Days, Tag, Length, FishCount, AgePhase) %>%
  distinct() %>% 
  group_by(Days, Fish_common) %>% 
  summarise(Average_Size_CBRUV = round(mean(Length), 1),
            overall_min_size = round(min(Length), 0),
            overall_max_size = round(max(Length), 0),
            sd_length = round(sd(Length, na.rm = TRUE),1),
            N_CBRUVS = n()) %>% 
  dplyr::select(Days, Fish_common, N_CBRUVS, Average_Size_CBRUV, overall_min_size, overall_max_size)  %>% 
  mutate(size_display_CBRUVS = sprintf("%.1f (%.0f–%.0f)", 
                               Average_Size_CBRUV, 
                               overall_min_size, 
                               overall_max_size))  -> CBRUV_Fish_Summary

full_join(RVC_Fish_Survey_subset, CBRUV_Fish_Summary, by = c("Days", "Fish_common")) %>% 
  filter(Days <= 14) %>% 
  write_csv("output_tables/CBRUV_RVC_Fish_Comparison.csv")
```


# GLMM for Cumulative Predation Prevalence
```{r}
PT1_clean %>% 
  filter(Species != "APAL" & Species != "ACER") %>% 
  filter(!Days == 0) %>%  # remove day 0 since zero predation
  dplyr::select(Days, Tray, Genotype, Coral, Cumulative) %>% 
  distinct() %>% 
  mutate(Cumulative = case_when(Cumulative == "Y" ~ 1,
                                Cumulative == "N" ~ 0)) %>% 
  mutate(Days = as.factor(Days)) -> predation_stats

glmm_predation <- glmmTMB::glmmTMB(Cumulative ~ Days + (1|Tray) + (1|Coral),
      family = binomial(link = "logit"), 
      data = predation_stats)

summary(glmm_predation)

Anova(glmm_predation, type = "II")

library(emmeans)
emm <- emmeans(glmm_predation, ~ Days)  # your actual days
pairs(emm, adjust = "tukey", type = "response")
summary(pairs(emm, adjust = "tukey", type = "response"), infer = TRUE)  # Adds confidence intervals
res <- summary(pairs(emm, adjust = "tukey", type = "response"), infer = TRUE)
as.data.frame(res)  # Shows all columns clearly

write_csv(as.data.frame(res), "output_tables/GLMM_predation_posthoc.csv") 
```


